Transformations of forecasts for evaluation in an epidemiological context

### 2022-11-29 main.tex
add citations

\title{Transformations of forecasts for evaluation in an epidemiological context}
  \author{Nikos I. Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,*}$,

  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$,

  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\

  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene & Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$,

  Johannes Bracher\thanks{Institute of Economic Theory and Statistics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$,

  Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$
  }

@@

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future. Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funk19_assessingperformancereal}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021}. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI). For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former is more commonly used for point forecasts and the latter is more appropriate for predictive distributions.

Infectious processes are commonly modelled around the concept of the reproduction number \citep{gosticPracticalConsiderationsMeasuring2020}, the average number of new cases generated by each case in the population.
In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time~\citep{dushoff2021_speedstrengthepidemic}.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellis21_challengescontrolcovid}.
If case numbers are growing exponentially (for an effective reproduction number greater than 1) and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
If one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.

>>>>

\title{Transformations of forecasts for evaluation in an epidemiological context}
  \author{Nikos I. Bosse\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{Centre for the Mathematical Modelling of Infectious Diseases, London, United Kingdom} $^{,*}$, 
  Sam Abbott\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$, 
  Anne Cori\thanks{MRC Centre for Outbreak Analysis and Modelling, Department of Infectious Disease Epidemiology, School of Public Health, Imperial College London, London, United Kingdom} $^{}$, \\
  Edwin van Leeuwen\thanks{Department of Infectious Disease Epidemiology, London School of Hygiene \& Tropical Medicine, London, United Kingdom} $^{,}$\thanks{UK Health Security Agency, London, United Kingdom} $^{}$, 
  Johannes Bracher\thanks{Institute of Economic Theory and Statistics, Karlsruhe Institute of Technology, Karlsruhe, Germany} $^{}$, 
  Sebastian Funk\footnotemark[1] $^{,}$\footnotemark[2]$ ^{}$}

@@

Probabilistic forecasts are usually evaluated using so-called proper scoring rules \citep{gneitingStrictlyProperScoring2007}, which return a numerical score as a function of the forecast and the observed data. Proper scoring rules are constructed such that forecasters are incentivised to report their true belief about the future. Examples of proper scoring rules that have been used to assess epidemiological forecasts are the Continuous Ranked Probability Score~\citep[CRPS,][]{gneitingStrictlyProperScoring2007} or its discrete equivalent, the Ranked Probability Score~\citep[RPS,][]{funkAssessingPerformanceRealtime2019}, and the Weighted Interval Score \citep[WIS,][formulas given in Section \ref{wis} in the SI]{bracherEvaluatingEpidemicForecasts2021}. The WIS is an approximation of the CRPS for predictive distributions represented by a set of predictive quantiles and is currently used to assess forecasts in COVID-19 forecast hubs in the US \citep{cramerCOVID19ForecastHub2020, cramerEvaluationIndividualEnsemble2021}, Europe \citep{sherrattPredictivePerformanceMultimodel2022}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, as well as the US Influenza Forecasting Hub \citep{CdcepiFlusightforecastdata2022}. The WIS and the CRPS measure the absolute distance between the predictive distribution and the observed value and can be understood as a generalisation of the absolute error to predictive distributions (see details in Sections \ref{wis} and \ref{crps} in the SI). For simplicity, we will use "absolute error" and "absolute distance" interchangeably, even though the former is more commonly used for point forecasts and the latter is more appropriate for predictive distributions.

Infectious processes are commonly modelled around the concept of the reproduction number \citep{gosticPracticalConsiderationsMeasuring2020}, the average number of new cases generated by each case in the population.
In the absence of changes in immunity, behaviour or other factors that may affect the intensity of transmission, the reproduction number would be expected to remain approximately constant.
In that case, the number of new infections in the population grows exponentially in time~\citep{dushoffSpeedStrengthEpidemic2021}.
This behaviour was observed, for example, early in the COVID-19 pandemic in many countries~\citep{pellisChallengesControlCOVID192021}.
If case numbers are growing exponentially (for an effective reproduction number greater than 1) and the true modelling task revolves around estimating the reproduction number or the corresponding growth rate, evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number) less than overprediction by the same amount, as errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.
If one is to assess the ability of forecasts to assess and forecast the underlying infection dynamics, it may be more appropriate to evaluate errors on the growth rate directly.
